---
title: "Practice visualizing the mpg dataset"
author: "Håkan Fridlund"
output: html_document
date: "2026-08-01"
lang: sv
header-includes:
  - "<style>body { hyphenate-character: none; hyphens: none; word-break: keep-all; }</style>"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(ggthemes)
```

# Introduction

In this report, I explore the `mpg`-dataset included in the **ggplot2** package.\
The dataset contains information about various car models, engine size, fuel economy, and vehicle classes.

I will visualize some key relationships in the data and have chosen to include code snippets to use as a reference for future visualizations.

```{r include=FALSE}
head(mpg)
?mpg
```

## Scatter plot

Shows the relationship between engine size and highway fuel consumption.

```{r}
ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm",
              se = FALSE, 
              color = "white") +
  labs(
    title = "Motorsize vs Miles per gallon on Highway ",
    x = "Motorsize (litres)",
    y = "Miles per gallon on Highway"
  ) 

```

## Box plot

Shows the distribution of fuel consumption on country roads for different types of operating modes.

```{r}
ggplot(mpg, aes(drv, hwy)) +
  geom_boxplot() +
  scale_x_discrete(
    labels = c(
      "4" = "4-wheel drive",
      "f" = "Front-wheel drive",
      "r" = "Rear-wheel drive"
    )
  ) +
  labs(
    title = "Fuel consumption per drivetrain",
    x = "Drivetrain",
    y = "Miles per gallon on Highway"
  )
```

## Histogram

Shows the distribution of fuel consumption.

```{r}
ggplot(mpg, aes(hwy)) +
  geom_histogram(bins = 20, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of fuel consumption",
    x = "Fuel consumption (Miles per gallon on Highway)",
    y = "Number of cars"
  )
```

## Faceted scatter plot

Shows the relationship between engine size and highway fuel consumption, broken down by class.

```{r}
ggplot(mpg, aes(displ, hwy, color=class)) +
  geom_point() +
  facet_wrap(~ class) +
  labs(
    title = "Motorsize vs Highway MPG by car class",
    x = "Motorsize i Litres",
    y = "Fuel consumption on highway"
  ) +
theme_minimal()

```
